---
title: "Lab 02 - Plastic waste"
author: "Ye Dam Yi"
date: "February 9, 2021"
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(ggsci)
```

```{r load-data, message=FALSE}
plastic_waste <- read_csv("data/plastic-waste.csv")
```

## Exercises

```{r}
ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap)) +
  geom_histogram(binwidth = 0.2)

plastic_waste %>%
  filter(plastic_waste_per_cap > 3.5)

```
### Exercise 1

Remove this text, and add your answer for Exercise 1 here.

```{r plastic-waste-continent}
ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap)) +
geom_histogram() +
  labs(
  x = "Plastic Waste Per Capita (kg/day)",
  y = "Count",
  title = "Plastic Waste Per Capiata (kg/day)",
  subtitle = "By Continent"
) + 
facet_wrap(~ continent)
```

### Exercise 2

```{r plastic-waste-density}
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     color = continent,
                     fill = continent)) +
  geom_density(alpha = .25)

```



### Exercise 3

We defined the color and fill of the curves by mapping aesthetics of the plot because we wanted the features to vary as the function of the variable continent. In contrast, we defined the alpha level as a characteristic of the plotting geom because we want the characteristic to apply to the graph universally.

### Exercise 4

```{r plastic-waste-boxplot}
ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_boxplot()
```

```{r plastic-waste-violin}
ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_violin() + 
  coord_flip() +
  labs(
    y = "Plastic Waste Per Capita (kg/day)",
    x = "Continent",
    title = "Plastic Waste Per Capita (kg/day)",
    subtitle = "By Continent"
  )
```
A comparison of the box plots and the violin plots shows that the violin plots reveal the density or the actual distribution of the data points of plastic waste per capital that the box plots do not reveal. In contrast, the box plots readily reveal the outliers in each of the plots as distinct dots far removed from the rest of the plots, which are subsumed in the violin plots and thus not really distinct. 

### Exercise 5

Remove this text, and add your answer for Exercise 5 here.

```{r plastic-waste-mismanaged}
ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap,
                                 y = mismanaged_plastic_waste_per_cap)) +
  geom_point() +
  labs(
    x= "Plastic Waste Per Capita (kg/day)",
    y = "Mismanaged Plastic Waste Per Cap",
    title = "Relationship between Plastic Waste Per Capita and Mismanaged Plastic Waste Per Capital"
    )
```
Countries seem to be somewhat polarized into two camps: plastic waste per capita is associated with high increases in mismanaged plastic waste per capita, or plastic waste per capita is associated with low increases in mismanaged plastic waste per capita. 

Many countries fall between the two trends. 

Interesting are two outliers, one that has the highest plastic waste per capita, and the other that has the highest mismanaged plastic waste per capital.


### Exercise 6

```{r plastic-waste-mismanaged-continent}
ggplot(data = plastic_waste, 
            aes(x = plastic_waste_per_cap, y = mismanaged_plastic_waste_per_cap,
                                 colour = continent)) +
  geom_point() +
  labs(
    x= "Plastic Waste Per Capita (kg/day)",
    y = "Mismanaged Plastic Waste Per Cap",
    title = "Relationship between Plastic Waste Per Capita and Mismanaged Plastic Waste Per Capital",
    subtitle = "By Continent"
    ) + 
  theme_dark() + theme(
    panel.background = element_rect(fill = "#2D2D2D"),
    legend.key = element_rect(fill = "#2D2D2D")) +
  scale_color_tron()
```

It seems that the two trends discussed earlier can capture certain continents. For example, Europe seems to follow the second trend discussed earlier whereby increases in plastic waste per capita relate to marginal increases in mismanaged plastic waste per capita. Asia follows both trends whereby some countries experience high or marginal increases in mismanaged plastic waste per capita as plastic waste per capita increases.
North America and South America follow a compromised, loosened pattern between the two trends. 

### Exercise 7

```{r plastic-waste-population-total}
ggplot(data = plastic_waste, 
            aes(x = plastic_waste_per_cap, y = total_pop
                                )) +
  geom_point() +
  labs(
    x= "Plastic Waste Per Capita (kg/day)",
    y = "Total Population",
    title = "Relationship between Plastic Waste Per Capita and Total Population"
    )  


```

```{r plastic-waste-population-coastal}
ggplot(data = plastic_waste, 
            aes(x = plastic_waste_per_cap, y = coastal_pop
                                )) +
  geom_point() +
  labs(
    x= "Plastic Waste Per Capita (kg/day)",
    y = "Coastal Population",
    title = "Relationship between Plastic Waste Per Capita and Coastal Population"
    )  
```

The relationships between plastic waste per capita and total population and between plastic waste per capita and coastal population seem to be similarly linear. The relationship between plastic waste per ccapita and total population, indicated by a more horizontal trend, seems closer to 0, and the relationship between plastic waste per capita and coastal population, indicated by a more vertical trend, seems closer to none as well. 

### Exercise 8


```{r recreate-viz}
# create a variable
plastic_waste$coastal_total_ratio <- plastic_waste$coastal_pop/plastic_waste$total_pop
# filter out an outlier 
plastic_waste %>%
  filter(plastic_waste_per_cap < 3.5) %>%
ggplot(aes(x = coastal_total_ratio, 
                y = plastic_waste_per_cap
                                )) +
  geom_point(aes(color = continent)) +
  labs(
    x= "Coastal population proportion (Coastal / total population)",
    y = "Plastic Waste Per Capita",
    title = "Plastic Waste vs. coastal population proportion",
    subtitle = "by continent"
    ) +
  stat_smooth(color = "black") + 
  scale_colour_viridis_d()
  
```
There is a zero relatinoship between coastal population proportion and plastic waste per capita as the global linear association between the two variable slightly increases, stays at zero change for a while, and then marginally declines as the coastal population proportion increases. This implies that there's no relationship between the extent to which coastal population constitues the total population in a country and the plastic waste per capita in that country.  

A couple of small patterns might be worth noting that do not pertain to the relationship between the two variables plotted above. It seems that plastic waste per capita differs for each content whereby Africa has the lowest and North America has the highest. Also, coastal population proportion seems to be capped at around .75 for countries in South America.

